---
title: "Relationship between types of food and negative impact on the environment"
execute: 
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
# Run the following commands to load the libraries.
library(ggplot2)
library(dplyr)
library(tidyverse)
```

```{r}
# I downloaded the dataset from Kaggle and loaded it into my R environment. Once the data is loaded, I create a plot to visualize the relationship between types of food and their negative impact on the environment.

#| label: plot
# Load the dataset
data <- read.csv("data/Food_Production.csv")
# Explore the dataset
# str(data)

# Data manipulation
# Depending on the structure of my dataset, I manipulated it to get it in the right format for plotting. Group by food type and get mean values of "Total emissions" and "The land used".
summary_data <- data %>%
  group_by(Food_product) %>%
  summarize(
    avg_emissions = mean(Total_emissions, na.rm = TRUE),
    avg_land_use = mean("Land use per 100g protein (m² per 100g protein)", na.rm = TRUE)
  )

# Plotting the data
# Use ggplot2 to create a plot showing the relationship between food types and their negative environmental impact.
ggplot(data, aes(x = Food_product, y = Total_emissions, fill = Food_product)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(
    title = "Environmental Impact of Different Food Types",
    x = "Type of Food",
    y = "Emissions"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
As the world’s population has expanded and gotten richer, the demand for food, energy and water has seen a rapid increase. Emissions from most plant based Foods are much lower than most animal based foods.

The following data was pulled from Kaggle. Then, the relevant data frames were joined to analyze the relationship between types of food and negative impact on the environment.

1. Looks like Beef is responsible for most of the greenhouse gas emissions.

2. Plant based foods are hardly visible in the graph indicating their significantly low carbon footprint.

3. Emissions from most plant based Foods are much lower than most animal based foods. 


As the world’s population has expanded and gotten richer, the demand for food, energy and water has seen a rapid increase. Not only has demand for all three increased, but they are also strongly interlinked: food production requires water and energy; traditional energy production demands water resources; agriculture provides a potential energy source. This data analysis focuses on the environmental impacts of food. Ensuring everyone in the world has access to a nutritious diet in a sustainable way is one of the greatest challenges we face.
